An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-26 VLDB2026/main.tex
change challenge

	Furthermore, current attribute filtering algorithms offer limited support for complex filtering conditions. For example, Filtered-DiskANN  supports single-attribute filtering, and in multi-attribute scenarios, it only supports OR conditions between attributes. 
	%The original implementations of 
	NHQ imposes strict constraints on the number of base and query attributes and only support AND conditions. While UNG is relatively flexible—supporting both AND and OR logic—it still lacks support for arbitrary Boolean expressions (e.g., combinations of AND and OR). The lack of flexible Boolean filtering remains a key challenge, limiting the practicality of hybrid query methods in real-world systems with diverse and complex query requirements.

@@

	Except for DSG, most range filtering algorithms require the dataset to be pre-sorted before building the index. Moreover, these methods do not support multi-attribute range queries. Enhancing these methods to support an arbitrary number of range filterable attributes remains an open research challenge.

	Our experiments show that graph compression methods often lead to poor query performance in RF-ANN search. To address this problem, subsequent algorithms usually optimize the index processing at the segment tree nodes. UNIFY newly proposes segment graphs and integrates multiple data structures to process queries. Although UNIFY performs slightly worse on small-range queries, its proposed index structure broadens the research direction.

		In our experiments, many range filtering algorithms perform worse than Faiss at high recall (0.95), showing that current algorithm designs still need improvement. We suggest future future range filtering algorithms include Faiss as a baseline.
		%In our experiments, we find that many range filtering algorithms perform worse than Faiss at high recall (0.95). This shows that the design of current algorithms still needs improvement. We suggest that future range filtering algorithms should include Faiss as a baseline.

@@

	Lastly, our findings show that the distribution and the number of attributes, as well as dataset properties (e.g., dimensionality and LID), significantly impact the performance of attribute filtering methods. Enhancing the robustness of these algorithms under diverse data conditions is another key challenge that warrants further investigation.

>>>>

	\textcolor{blue}{In addition, existing attribute filtering algorithms face two major limitations: strict requirements on attribute format and limited support for complex filtering conditions. For example, NHQ organizes attributes by column, which is suitable for uniformly structured data, but performs poorly in scenarios where the number of attributes varies across data points, resulting in poor scalability. Furthermore, NHQ only supports multi-attribute filtering with conjunctions ("AND" conditions). UNG is relatively more flexible in this regard, as it supports both "AND" and "OR" logic; however, it still cannot handle more complex Boolean expressions involving combinations of conjunctions and disjunctions.
	In contrast, FilterDiskANN imposes no restrictions on the number of attributes per data point, offering better scalability. However, it currently supports only single-attribute filtering and disjunctions ("OR" conditions) across multiple attributes. Enhancing its capability to support more flexible Boolean combinations would effectively overcome the limitations of current approaches and significantly improve its practicality for diverse and complex query requirements.
		}

	\textcolor{blue}{
	Besides, our research findings indicate that the distribution and selectivity of attributes significantly impact the performance of attribute filtering methods. Therefore, when designing algorithms, efforts should be made to minimize the coupling with specific attribute value distributions, in order to enhance the robustness and generalization ability of the algorithms under varying data conditions.}

%	Furthermore, current attribute filtering algorithms offer limited support for complex filtering conditions. For example, Filtered-DiskANN  supports single-attribute filtering, and in multi-attribute scenarios, it only supports OR conditions between attributes. 
%	%The original implementations of 
%	NHQ imposes strict constraints on the number of base and query attributes and only support AND conditions. While UNG is relatively flexible—supporting both AND and OR logic—it still lacks support for arbitrary Boolean expressions (e.g., combinations of AND and OR). The lack of flexible Boolean filtering remains a key challenge, limiting the practicality of hybrid query methods in real-world systems with diverse and complex query requirements.

@@

	Except for DSG and UNIFY, most range filtering algorithms require the dataset to be pre-sorted before building the index. Moreover, these methods do not support multi-attribute range queries. Enhancing these methods to support an arbitrary number of range filterable attributes remains an open research challenge.

%	Our experiments show that graph compression methods often lead to poor query performance in RF-ANN search. To address this problem, subsequent algorithms usually optimize the index processing at the segment tree nodes. UNIFY newly proposes segment graphs and integrates multiple data structures to process queries. Although UNIFY performs slightly worse on small-range queries, its proposed index structure broadens the research direction.

\textcolor{blue}{
Experimental results show that different algorithms exhibit varying strengths under different query ranges. Therefore, designing a general range query algorithm that can adapt to diverse query conditions holds significant research value. For example, UNIFY demonstrates strong generality by dynamically selecting execution strategies based on the query range. Building on this idea, further exploration of more adaptive and high-performance range query algorithms is a promising direction for future research.}

		In our experiments, many range filtering algorithms perform worse than Faiss at high recall (0.95), showing that current algorithm designs still need improvement. We suggest future future range filtering algorithms include Faiss as a baseline.

@@

	\textcolor{blue}{
	Lastly, existing algorithms suffer from significant performance degradation on out-of-distribution (OOD) datasets. Exploring how to improve these algorithms to better adapt to more complex query conditions is a worthwhile direction for further research.}

%	our findings show that the distribution and the number of attributes, as well as dataset properties (e.g., dimensionality and LID), significantly impact the performance of attribute filtering methods. Enhancing the robustness of these algorithms under diverse data conditions is another key challenge that warrants further investigation.
%
